---
title: "GAL settling chapter 1"
output: pdf_document
date: "2024-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)

```{r setup, include=FALSE}
setwd("C:/git/SSEE/SSEE Project")  # Force set the working directory
knitr::opts_knit$set(root.dir = "C:/git/SSEE/SSEE Project")

getwd()
```

```{r}

#galdata=read.csv("GAL settling phenotyping 2024 T3 14.csv")

galdata=read.csv("./Original Data/GAL settling phenotyping 2024 T3 14.csv")

#Order treatments so that all graphs look the same, this is just the first step to doing that though. 
galdata$Treatment <- factor(galdata$Treatment, levels = c("YPD","YPDSET","GAL", "GALSET"))

galdata=GAL.settling.phenotyping.2024.T3.14

```


```{r}
library(ggplot2)
library(dplyr)
library(patchwork)

# Define the transfer times you are interested in
transfer_times <- c(1, 4, 8, 12, 16, 20, 24, 28)

# Function to filter, clean, and calculate the mean
filter_clean_calc_mean <- function(data, treatment, transfer) {
  filtered_data <- filter(data, Treatment == treatment, Transfer == transfer) %>% na.omit()
  avg_value <- mean(filtered_data$Number, na.rm = TRUE)
  return(list(data = filtered_data, avg = avg_value))
}

# Process GALSET and GAL treatments
galset_results <- lapply(transfer_times, function(t) filter_clean_calc_mean(galdata, "GALSET", t))
gal_results <- lapply(c(1, 4), function(t) filter_clean_calc_mean(galdata, "GAL", t))

# Create a function to generate histograms
generate_histogram <- function(data, avg, title_prefix, fill_color, x_limit = 204, y_limit = 200) {
  ggplot(data, aes(x = Number)) +
    geom_histogram(binwidth = 1, fill = fill_color, color = "black", alpha = 0.7) +
    ggtitle(paste(title_prefix, "Settling time", unique(data$Transfer), "avg", round(avg, 2))) +
    xlab("Clump size") +
    ylab("Frequency") +
    theme_minimal() +
    xlim(0, x_limit) +
    ylim(0, y_limit)
}

# Generate plots for GALSET
galset_plots <- mapply(generate_histogram, 
                        lapply(galset_results, `[[`, "data"), 
                        lapply(galset_results, `[[`, "avg"), 
                        MoreArgs = list(title_prefix = "GALSET", fill_color = "#4848FF"), 
                        SIMPLIFY = FALSE)

# Generate plots for GAL
gal_plots <- mapply(generate_histogram, 
                    lapply(gal_results, `[[`, "data"), 
                    lapply(gal_results, `[[`, "avg"), 
                    MoreArgs = list(title_prefix = "GAL", fill_color = "#ADD8E6", x_limit = 30, y_limit = 300), 
                    SIMPLIFY = FALSE)

# Combine and display plots
combined_plot <- (gal_plots[[1]] / gal_plots[[2]] / galset_plots[[1]] / galset_plots[[2]]) + plot_layout(ncol = 2)
combined_plot

plot2 <- wrap_plots(galset_plots, ncol = 2)
plot2

plot3 <- galset_plots[[4]] / galset_plots[[6]] + plot_layout(ncol = 2)
plot3

plot4 <- galset_plots[[5]] / galset_plots[[6]] / galset_plots[[7]] / galset_plots[[8]] + plot_layout(ncol = 2)
plot4

plot5 <- galset_plots[[1]] / galset_plots[[4]] / galset_plots[[8]] + plot_layout(ncol = 3)
plot5




```
```{r}
library(dplyr)
library(ggplot2)
library(vegan)  # For Shannon and Simpson diversity calculations
library(patchwork)
library(tidyr)

# Define transfer times and treatments
transfer_times <- c(1, 4, 8, 12, 16, 20, 24, 28)
treatments <- c("GALSET", "GAL", "YPDSET", "YPD")

# Function to calculate diversity metrics
calculate_diversity <- function(data) {
  if (nrow(data) == 0) return(c(Shannon = NA, Simpson = NA, Variance = NA))  # Avoid errors
  
  freq_table <- table(data$Number)  # Frequency of each clump size
  proportions <- freq_table / sum(freq_table)  # Convert to proportions
  
  shannon_index <- -sum(proportions * log(proportions), na.rm = TRUE)  # Shannon Index
  simpson_index <- sum(proportions^2, na.rm = TRUE)  # Simpson Index
  variance <- var(data$Number, na.rm = TRUE)  # Variance
  
  return(c(Shannon = shannon_index, Simpson = simpson_index, Variance = variance))
}

# Collect diversity data for all treatments
diversity_results <- list()

for (treatment in treatments) {
  for (t in transfer_times) {
    filtered_data <- filter(galdata, Treatment == treatment, Transfer == t) %>% na.omit()
    
    if (nrow(filtered_data) > 0) {  # Only add non-empty data
      diversity_metrics <- calculate_diversity(filtered_data)
      diversity_results <- append(diversity_results, list(
        data.frame(Treatment = treatment, Transfer = t,
                   Shannon = diversity_metrics["Shannon"],
                   Simpson = diversity_metrics["Simpson"],
                   Variance = diversity_metrics["Variance"])
      ))
    }
  }
}

# Combine results into a single dataframe
diversity_df <- bind_rows(diversity_results)

# Check if the data is non-empty
print(diversity_df)  # Debugging step: make sure data exists

# Reshape data for ggplot (long format)
diversity_long <- diversity_df %>%
  pivot_longer(cols = c(Shannon, Simpson, Variance), names_to = "Metric", values_to = "Value") %>%
  drop_na()  # Remove any NA values to prevent blank plots

# Debug: Check reshaped data
print(diversity_long)

# Define color scheme for treatments
color_scheme <- c("GALSET" = "#4848FF", "GAL" = "#ADD8E6", "YPDSET" = "#006400", "YPD" = "#90EE90")

# Custom plot function with y-axis limits auto-adjusted
generate_diversity_plot <- function(metric_name, y_label, y_scale="linear") {
  filtered_data <- filter(diversity_long, Metric == metric_name)
  
  if (nrow(filtered_data) == 0) {
    message(paste("No data available for", metric_name))
    return(ggplot() + ggtitle(paste("No data for", metric_name)))
  }
  
  plot <- ggplot(filtered_data, aes(x = Transfer, y = Value, color = Treatment, group = Treatment)) +
    geom_line(size = 1.2) + geom_point(size = 4) +
    ggtitle(paste(metric_name, "Over Time")) +
    xlab("Transfer Time") + ylab(y_label) +
    theme_minimal(base_size = 16) +
    theme(
      legend.position = "none",  # Remove legends from individual plots
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text = element_text(size = 14),
      axis.title = element_text(size = 16),
      plot.margin = margin(20, 20, 20, 20)
    ) +
    scale_color_manual(values = color_scheme) +
    scale_x_continuous(breaks = transfer_times)  # Ensure x-axis labels are correct
  
  if (y_scale == "log") {
    plot <- plot + scale_y_log10()
  } else {
    plot <- plot + scale_y_continuous()
  }
  
  return(plot)
}

# Generate individual plots
shannon_plot <- generate_diversity_plot("Shannon", "Shannon Diversity Index")
simpson_plot <- generate_diversity_plot("Simpson", "Simpson Index (Lower = More Diversity)")
variance_plot <- generate_diversity_plot("Variance", "Variance in Clump Size", y_scale="log")

# Arrange all plots in a vertical layout and collect the legend
final_plot <- (shannon_plot / simpson_plot / variance_plot) + 
  plot_layout(ncol = 1, guides = "collect") & 
  theme(legend.position = "bottom")  # Single legend at bottom

# Display final plot
final_plot


```


```{r}
library(dplyr)
library(ggplot2)
library(patchwork)

# Define the transfer times you are interested in
transfer_times <- c(1, 4, 8, 12, 16, 20, 24, 28)

# Function to filter, clean, and calculate the mean
filter_clean_calc_mean <- function(data, treatment, transfer, reps = c(1,2,3,4)) {
  filtered_data <- filter(data, Treatment == treatment, Transfer == transfer, Replicate %in% reps) %>% na.omit()
  avg_value <- mean(filtered_data$Number, na.rm = TRUE)
  return(list(data = filtered_data, avg = avg_value))
}

# Process YPDSET and YPD treatments
ypdset_results <- lapply(transfer_times, function(t) filter_clean_calc_mean(galdata, "YPDSET", t))
ypd_results <- lapply(c(1, 4), function(t) filter_clean_calc_mean(galdata, "YPD", t))

# Create a function to generate histograms
generate_histogram <- function(data, avg, title_prefix, fill_color, x_limit = 30, y_limit = 200) {
  ggplot(data, aes(x = Number)) +
    geom_histogram(binwidth = 1, fill = fill_color, color = "black", alpha = 0.7) +
    ggtitle(paste(title_prefix, "Settling time", unique(data$Transfer), "avg", round(avg, 2))) +
    xlab("Clump size") +
    ylab("Frequency") +
    theme_minimal() +
    xlim(0, x_limit) +
    ylim(0, y_limit)
}

# Generate plots for YPDSET
ypdset_plots <- mapply(generate_histogram, 
                        lapply(ypdset_results, `[[`, "data"), 
                        lapply(ypdset_results, `[[`, "avg"), 
                        MoreArgs = list(title_prefix = "YPDSET", fill_color = "#006400"), 
                        SIMPLIFY = FALSE)

# Generate plots for YPD
ypd_plots <- mapply(generate_histogram, 
                    lapply(ypd_results, `[[`, "data"), 
                    lapply(ypd_results, `[[`, "avg"), 
                    MoreArgs = list(title_prefix = "YPD", fill_color = "#90EE90", x_limit = 20, y_limit = 400), 
                    SIMPLIFY = FALSE)

# Combine and display plots
combined_plot <- (ypd_plots[[1]] / ypd_plots[[2]] / ypdset_plots[[1]] / ypdset_plots[[2]]) + plot_layout(ncol = 2)
combined_plot

# Manually select and display plots the same way as GALSET
YPDsetplotCombo <- ypdset_plots[[1]] / ypdset_plots[[2]] / ypdset_plots[[3]] / ypdset_plots[[4]] + plot_layout(ncol = 2)
YPDsetplotCombo

YPDSET <- ypdset_plots[[5]] / ypdset_plots[[6]] / ypdset_plots[[7]] / ypdset_plots[[8]] + plot_layout(ncol = 2)
YPDSET




```


```{r}

# Loop over treatments to generate graphs with customized axes
for (treatment in unique(galdata$Treatment)) {
  treatment_summary <- galdata_summary %>% filter(Treatment == treatment)
  
  # Create the plot with custom x-axis ticks
  plot <- ggplot(treatment_summary, aes(x = Transfer, y = Total_Cells, color = Binned_Number, group = Binned_Number)) +
    geom_line(size = 1.2) +
    geom_point(size = 2) +
    scale_x_continuous(breaks = custom_x_ticks) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    labs(title = paste(treatment, "- Total Cells in Clumps Across Transfers"),
         x = "Transfer",
         y = "Total Cells in Clumps",
         color = "Clump Size Bin") +
    theme_minimal() +
    theme(legend.position = "right")
  
  # Print the plot
  print(plot)
}

multipaneled_plot_fixed_y <- ggplot(galdata_summary, aes(x = Transfer, y = Total_Cells, color = Binned_Number, group = Binned_Number)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  facet_wrap(~ Treatment, scales = "fixed") +  # Use fixed scales for consistent y-axis
  scale_x_continuous(breaks = custom_x_ticks) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(title = "Total Cells in Clumps Across Transfers by Treatment (Fixed Y-Axis)",
       x = "Transfer",
       y = "Total Cells in Clumps",
       color = "Clump Size Bin") +
  theme_minimal() +
  theme(legend.position = "right")

# Print the multipaneled plot
print(multipaneled_plot_fixed_y)


```


```{R}
#This is a model for the segements using the total cells "galdata_summary" 
# Load necessary libraries
library(tidyverse)

# Function to fit a model and show its summary
fit_and_summarize <- function(data, treatment, bin, segment_points) {
  # Filter the data for the given treatment and bin
  data_subset <- data %>%
    filter(Treatment == treatment, Binned_Number == bin)
  
  # Add a segment variable based on segment_points
  data_subset <- data_subset %>%
    mutate(Segment = case_when(
      Transfer <= segment_points[1] ~ "Segment1",
      Transfer > segment_points[1] & Transfer <= segment_points[2] ~ "Segment2",
      TRUE ~ "Segment3" # Optional for bins with more than two segments
    ))
  
  # Fit the model
  interaction_model <- lm(Total_Cells ~ Transfer * Segment, data = data_subset)
  
  # Return the model summary
  summary(interaction_model)
}

# Example Usage
# Define segmentation points for the specific bin
segment_points_5_9 <- c(8, 28)  # For Bin 5-9
segment_points_10_19 <- c(12, 28) # For Bin 10-19

# Call the function for a specific treatment and bin
fit_and_summarize(galdata_summary, treatment = "GALSET", bin = "5-9", segment_points = segment_points_5_9)


fit_and_summarize(galdata_summary, treatment = "GALSET", bin = "10-19", segment_points = c(12, 28))

fit_and_summarize(galdata_summary, treatment = "GALSET", bin = "20-29", segment_points = c(16, 28))

fit_and_summarize(galdata_summary, treatment = "YPDSET", bin = "10-19", segment_points = c(20, 28))

fit_and_summarize(galdata_summary, treatment = "YPDSET", bin = "20-29", segment_points = c(20, 28))


```

```{r}

# Custom color palette to match distinct bins
custom_colors <- c("#08306B", "#2171B5", "#4292C6", "#6BAED6", "#9ECAE1", 
                   "#C6DBEF", "#FDD0A2", "#F16913", "#D94801", 
                   "#A50F15", "#67000D")

# Plot the stacked bar chart
ggplot(data_summary, aes(x = factor(Transfer), y = Proportion, fill = Clump_Size_Bin)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Treatment, scales = "free_x") +
  scale_fill_manual(values = custom_colors, name = "Clump Size Bin") +
  labs(title = "Clump Size Proportions Across Transfers by Treatment",
       x = "Transfer",
       y = "Proportion of Total Cells") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right",
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

```



```{r}

custom_colors <- c("#08306B", "#2171B5", "#4292C6", "#6BAED6", "#9ECAE1", 
                   "#C6DBEF", "#FDD0A2", "#F16913", "#D94801", 
                   "#A50F15", "#67000D")

# Plot the stacked bar chart
ggplot(Total_Cells, aes(x = factor(Transfer), y = Proportion, fill = Binned_Number)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~Treatment, scales = "free_x") +
  scale_fill_manual(values = custom_colors, name = "Clump Size Bin") +
  labs(title = "Total Cells Proportions Across Transfers by Treatment",
       x = "Transfer",
       y = "Proportion of Total Cells") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right",
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))



```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a custom color palette with one color per treatment
color_palette <- c(
  "YPD" = "#90EE90",   # Light green
  "YPDSET" = "#006400",  # Dark green
  "GAL" = "#ADD8E6",   # Light blue
  "GALSET" = "#4848FF"  # Dark blue
)

# Assuming 'galdata' is your dataset
# Filtering for Transfer 1 and 4
transfer_filtered <- galdata %>% filter(Transfer %in% c(1, 4, 8, 12, 16, 20, 24, 28))

# Grouping by Treatment, Replicate, Settling status, and Transfer to calculate the average of the 'Number' column
avg_transfer <- transfer_filtered %>%
  group_by(Treatment, Replicate, Settling, Transfer) %>%
  summarise(Average = mean(Number, na.rm = TRUE)) %>%
  ungroup()

# Create individual plots for each treatment and combine them side by side
ggplot(avg_transfer, aes(x = Transfer, y = Average, color = Treatment)) +
  geom_line(aes(group = Replicate), size = 1) +
  geom_point(aes(group = Replicate), size = 2) +
  facet_wrap(~ Treatment, ncol = 4) +
  scale_color_manual(values = color_palette) +  # Applying the custom color palette
  scale_x_continuous(breaks = c(1, 4, 8, 12, 16, 20, 24, 28)) +  # Ensure x-axis only includes Transfer 1 and 4
  labs(title = "Change in Multicellularity Between Transfer 1 and 28",
       x = "Transfer",
       y = "Average Number per Group") +
  theme_minimal() +
  theme(legend.position = "none")  # Hide the legend since each plot is labeled




``` 

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a custom color palette with one color per treatment
color_palette <- c(
  "YPD" = "#90EE90",   # Light green
  "YPDSET" = "#006400",  # Dark green
  "GAL" = "#ADD8E6"    # Light blue
)

# Assuming 'galdata' is your dataset
# Filtering for specific Transfers and excluding "GALSET" in Treatment
transfer_filtered <- galdata %>%
  filter(Transfer %in% c(1, 4, 8, 12, 16, 20, 24, 28) & Treatment %in% c("YPD", "YPDSET", "GAL"))

# Grouping by Treatment, Replicate, Settling status, and Transfer to calculate the average of the 'Number' column
avg_transfer <- transfer_filtered %>%
  group_by(Treatment, Replicate, Settling, Transfer) %>%
  summarise(Average = mean(Number, na.rm = TRUE)) %>%
  ungroup()

# Create individual plots for each treatment and combine them side by side
ggplot(avg_transfer, aes(x = Transfer, y = Average, color = Treatment)) +
  geom_line(aes(group = Replicate), size = 1) +
  geom_point(aes(group = Replicate), size = 2) +
  facet_wrap(~ Treatment, ncol = 3) +  # Adjusting columns since only 3 treatments remain
  scale_color_manual(values = color_palette) +  # Applying the custom color palette
  scale_x_continuous(breaks = c(1, 4, 8, 12, 16, 20, 24, 28)) +  # Ensure x-axis includes only specified transfers
  labs(title = "Change in Multicellularity Across Transfers",
       x = "Transfer",
       y = "Average Number per Group") +
  theme_minimal() +
  theme(legend.position = "none")  # Hide the legend since each plot is labeled





``` 

```{r}



# Separate data for "Settled" and "Non-Settled" treatments
settled_data <- ODs %>% filter(Treatment %in% c("GALSET Settled", "YPDSET Settled"))
non_settled_data <- ODs %>% filter(!Treatment %in% c("GALSET Settled", "YPDSET Settled"))

# Summarize data
summarize_ODs <- function(Data) {
  ODs %>%
    group_by(Treatment, Date) %>%
    summarise(
      Mean_OD = mean(OD, na.rm = TRUE),
      SEM_OD = sd(OD, na.rm = TRUE) / sqrt(n())
    ) %>%
    ungroup()
}

summary_settled <- summarize_ODs(settled_data)
summary_non_settled <- summarize_ODs(non_settled_data)

# Plot Non-Settled
plot_non_settled <- ggplot(summary_non_settled, aes(x = Date, y = Mean_OD, color = Treatment, group = Treatment)) +
  geom_line(size = 0.8) +
  geom_point(size = 1.5) +
  geom_ribbon(aes(ymin = Mean_OD - SEM_OD, ymax = Mean_OD + SEM_OD, fill = Treatment), alpha = 0.2) +
  labs(title = "Non-Settled Treatments: OD Measurements Through Transfers",
       subtitle = "Mean OD with SEM for GAL, GALSET, YPD, and YPDSET",
       x = "Transfer Date",
       y = "Mean OD ± SEM",
       color = "Treatment",
       fill = "Treatment") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

# Plot Settled
plot_settled <- ggplot(summary_settled, aes(x = Date, y = Mean_OD, color = Treatment, group = Treatment)) +
  geom_line(size = 0.8) +
  geom_point(size = 1.5) +
  geom_ribbon(aes(ymin = Mean_OD - SEM_OD, ymax = Mean_OD + SEM_OD, fill = Treatment), alpha = 0.2) +
  labs(title = "Settled Treatments: OD Measurements Through Transfers",
       subtitle = "Mean OD with SEM for GALSET Settled and YPDSET Settled",
       x = "Transfer Date",
       y = "Mean OD ± SEM",
       color = "Treatment",
       fill = "Treatment") +
  scale_color_brewer(palette = "Set3") +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

# Print plots
print(plot_non_settled)
print(plot_settled)
```

```{r}


# Clean column names
names(ODs) <- trimws(names(ODs))

# Normalize the Treatment column
ODs$Treatment <- trimws(ODs$Treatment)

# Filter out Settled treatments
filtered_ODs <- ODs %>%
  filter(!Treatment %in% c("GALSET Settled", "YPDSET Settled"))

# Convert Date column to proper Date format
filtered_ODs$Date <- as.Date(filtered_ODs$Date, format = "%Y-%m-%d")

# Remove rows with invalid dates
filtered_ODs <- filtered_ODs %>% filter(!is.na(Date))

# Summarize the data and handle missing or insufficient data
summary_ODs <- filtered_ODs %>%
  group_by(Treatment, Date) %>%
  summarise(
    Mean_OD = mean(OD, na.rm = TRUE),
    SEM_OD = ifelse(n() > 1, sd(OD, na.rm = TRUE) / sqrt(n()), NA_real_)
  ) %>%
  filter(!is.na(Mean_OD)) %>%
  ungroup()

# Check for empty summary_data
if (nrow(summary_ODs) == 0) {
  stop("The summary_data dataframe is empty. Please check your data and filtering steps.")
}

# Define custom colors for each treatment
custom_colors <- c(
  "YPD" = "#32CD32",     # YPD.B
  "YPDSET" = "#008000",  # YPDSET.B
  "GAL" = "#87CEEB",     # GAL.B
  "GALSET" = "#4169E1"   # GALSET.B
)

# Ensure Treatment is a factor for consistent ordering
summary_ODs$Treatment <- factor(summary_ODs$Treatment, levels = c("YPD", "YPDSET", "GAL", "GALSET"))

# Create the graph
ggplot(summary_ODs, aes(x = Date, y = Mean_OD, color = Treatment, group = Treatment)) +
  geom_line(size = 0.8) +
  geom_point(size = 1.5) +
  geom_ribbon(aes(ymin = Mean_OD - SEM_OD, ymax = Mean_OD + SEM_OD, fill = Treatment), alpha = 0.2) +
  facet_wrap(~ Treatment, ncol = 4, scales = "fixed") +
  labs(title = "OD Measurements for Non-Settled Treatments",
       subtitle = "Mean OD with SEM for GAL, GALSET, YPD, and YPDSET",
       x = "Transfer Date",
       y = "Mean OD ± SEM",
       color = "Treatment",
       fill = "Treatment") +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  scale_x_date(
    date_breaks = "1 week",    # Set date breaks to 1 week
    date_labels = "%b %d"      # Format date labels as "Jan 01"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )



```
```{r}
library(ggplot2)
library(dplyr)

# Clean column names
names(ODs) <- trimws(names(ODs))

# Normalize the Treatment column
ODs$Treatment <- trimws(ODs$Treatment)

# Convert Date column to proper Date format
ODs$Date <- as.Date(ODs$Date, format = "%Y-%m-%d")

# Remove rows with invalid dates
ODs <- ODs %>% filter(!is.na(Date))

# Summarize the data
summary_ODs <- ODs %>%
  group_by(Treatment, Date) %>%
  summarise(
    Mean_OD = mean(OD, na.rm = TRUE),
    SEM_OD = ifelse(n() > 1, sd(OD, na.rm = TRUE) / sqrt(n()), NA_real_),
    .groups = "drop"
  ) %>%
  filter(!is.na(Mean_OD))

# Check for empty summary data
if (nrow(summary_ODs) == 0) {
  stop("The summary_data dataframe is empty. Please check your data and filtering steps.")
}

# Define custom colors for each treatment, including settled treatments
custom_colors <- c(
  "YPD" = "#32CD32",            # Green
  "YPDSET" = "#008000",         # Dark Green
  "GAL" = "#87CEEB",            # Light Blue
  "GALSET" = "#4169E1",         # Royal Blue
  "GALSET Settled" = "#00008B", # Dark Blue
  "YPDSET Settled" = "#006400"  # Darker Green
)

# Ensure Treatment is a factor for consistent ordering
summary_ODs$Treatment <- factor(summary_ODs$Treatment, levels = c("YPD", "YPDSET", "GAL", "GALSET", "YPDSET Settled", "GALSET Settled"))

# Create the graph with facets for each treatment
ggplot(summary_ODs, aes(x = Date, y = Mean_OD, color = Treatment, group = Treatment)) +
  geom_line(size = 0.8) +
  geom_point(size = 1.5) +
  geom_ribbon(aes(ymin = Mean_OD - SEM_OD, ymax = Mean_OD + SEM_OD, fill = Treatment), alpha = 0.2) +
  facet_wrap(~ Treatment, ncol = 6, scales = "fixed") +  # Ensure all treatments are next to each other
  labs(title = "OD Measurements for All Treatments",
       subtitle = "Mean OD with SEM for GAL, GALSET, YPD, YPDSET, and Settled Treatments",
       x = "Transfer Date",
       y = "Mean OD ± SEM",
       color = "Treatment",
       fill = "Treatment") +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%b %d"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )






```



```{r}
library(ggplot2)
library(dplyr)

# Ensure Date is in proper date format in the raw data
filtered_ODs <- filtered_ODs %>%
  mutate(
    Date = as.Date(Date, format = "%m/%d/%Y"),
    Treatment = factor(Treatment, levels = c("YPD", "YPDSET", "GAL", "GALSET")) # Set treatment order
  )

# Summarize the data for plotting
summary_ODs <- filtered_ODs %>%
  group_by(Treatment, Date) %>%
  summarise(
    Mean_OD = mean(OD, na.rm = TRUE),
    SEM_OD = sd(OD, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Fit the linear model to generate predictions
model <- lm(OD ~ Date * Treatment, data = filtered_ODs, na.action = na.exclude)

# Check model summary for debugging
summary(model)

# Generate predictions for the model
filtered_ODs <- filtered_ODs %>%
  mutate(Predicted = predict(model))

# Summarize model predictions for plotting
OD_predictions <- filtered_ODs %>%
  group_by(Treatment, Date) %>%
  summarise(
    Mean_Predicted = mean(Predicted, na.rm = TRUE),
    .groups = "drop"
  )

# Define custom colors for each treatment
custom_colors <- c(
  "YPD" = "#32CD32",     # YPD.B
  "YPDSET" = "#008000",  # YPDSET.B
  "GAL" = "#87CEEB",     # GAL.B
  "GALSET" = "#4169E1"   # GALSET.B
)

# Plot the data
ggplot(summary_ODs, aes(x = Date, y = Mean_OD, color = Treatment, group = Treatment)) +
  geom_line(size = 0.8) + # Observed mean OD
  geom_point(size = 1.5) + # Observed data points
  geom_ribbon(aes(ymin = Mean_OD - SEM_OD, ymax = Mean_OD + SEM_OD, fill = Treatment), alpha = 0.2) + # SEM shading
  geom_line(data = OD_predictions, aes(x = Date, y = Mean_Predicted, color = Treatment), 
            linetype = "solid", size = 1.2, alpha = 0.8) + # Enhanced model predictions
  facet_wrap(~ Treatment, ncol = 4, scales = "fixed") + # Panels for each treatment
  labs(
    title = "OD Measurements for Non-Settled Treatments",
    subtitle = "Mean OD with SEM and Model Predictions for GAL, GALSET, YPD, and YPDSET",
    x = "Transfer Date",
    y = "Mean OD ± SEM",
    color = "Treatment",
    fill = "Treatment"
  ) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12), # Larger facet labels for clarity
    legend.position = "bottom"
  )





```


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Custom colors for the treatments
custom_colors <- c(
  "YPD" = "#32CD32",     # YPD.B
  "YPDSET" = "#008000",  # YPDSET.B
  "GAL" = "#87CEEB",     # GAL.B
  "GALSET" = "#4169E1"   # GALSET.B
)

# Compute summary statistics with SEM based on exactly 8 replicates
summary_data <- galdata %>%
  group_by(Treatment, Transfer) %>%
  summarize(
    mean = mean(Number, na.rm = TRUE),
    sem = sd(Number, na.rm = TRUE) / sqrt(8),  # Calculate SEM using 8 replicates
    lower = pmax(mean - sem, 0),              # Ensure lower bound is non-negative
    upper = mean + sem,                       # Upper bound
    .groups = "drop"                          # Ungroup after summarization
  )
summary_data$Treatment <- factor(summary_data$Treatment, levels = c("GAL", "GALSET", "YPD", "YPDSET"))
# Generate the plot with faceted panels and shared scales
ggplot(summary_data, aes(x = Transfer, y = mean, color = Treatment, fill = Treatment)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  scale_color_manual(values = custom_colors) +  # Apply custom colors for lines
  scale_fill_manual(values = custom_colors) +   # Apply custom colors for ribbons
  labs(y = "Average Number", x = "Transfer", title = "Variation of Number across Replicates (SEM)") +
  facet_wrap(~ Treatment, scales = "fixed") +  # Panels share the same scale
  theme_minimal() +
  theme(
    legend.position = "none",  # Remove the legend since treatments are in separate panels
    strip.text = element_text(size = 12, face = "bold"),  # Style the panel titles
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

# Calculate averages for each treatment and transfer
averages <- data %>%
  group_by(Treatment, Transfer) %>%   # Group by Treatment and Transfer
  summarise(Average = mean(Number, na.rm = TRUE),  # Mean of the 'Number' column
            SD = sd(Number, na.rm = TRUE),        # Standard deviation
            N = n())                             # Number of observations

# Display the averaged data
print(averages)

# Perform ANOVA and Tukey's HSD for each transfer
anova_results <- data %>%
  group_by(Transfer) %>%
  group_split() %>%
  lapply(function(df) {
    transfer_num <- unique(df$Transfer)
    anova_model <- aov(Number ~ Treatment, data = df)
    tukey <- TukeyHSD(anova_model)  # Perform Tukey's HSD post hoc test
    list(Transfer = transfer_num, ANOVA = summary(anova_model), TukeyHSD = tukey)
  })

# Display ANOVA and Tukey's HSD results for each transfer
for (result in anova_results) {
  cat("\nTransfer:", result$Transfer, "\n")
  print(result$ANOVA)
  cat("\nTukey's HSD Results:\n")
  print(result$TukeyHSD)
}




```

```{r}
##### CODE BELOW FOR LINEAR MODEL COMPARING REPLICATE MEANS ACROSS THE EXPERIMENT BETWEEN TRANSFERS AND THROUGHT THE TREATMENTS. 
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Step 1: Aggregate the real data
avg_data <- galdata %>%
  group_by(Replicate, Transfer, Treatment) %>%
  summarise(Average_Number = mean(Number, na.rm = TRUE), .groups = 'drop')

# Step 2: Fit the linear model
AVGmodel <- lm(Average_Number ~ Treatment * Transfer, data = avg_data)

summary(AVGmodel)

# Step 3: Create predictions for visualization
# Generate a new dataset for predictions
predict_data <- avg_data %>%
  distinct(Treatment, Transfer) %>%
  mutate(Average_Number_Predicted = predict(AVGmodel, newdata = .))

# Step 4: Visualize the data
ggplot() +
  # Plot real data points
  geom_point(data = avg_data, aes(x = Transfer, y = Average_Number, color = Treatment), size = 3, alpha = 0.6) +
  # Plot predicted values
  geom_line(data = predict_data, aes(x = Transfer, y = Average_Number_Predicted, color = Treatment), size = 1.2) +
  # Customize the plot
  labs(
    title = "Real Data and Model Predictions Over Transfers",
    x = "Transfer",
    y = "Average Number",
    color = "Treatment"
  ) +
  theme_minimal()

# Load necessary library
library(emmeans)

# Step 1: Fit the model (if not already done)
model <- lm(Average_Number ~ Treatment * Transfer, data = avg_data)

# Step 2: Perform post hoc comparisons accounting for interactions
# Obtain estimated marginal means for Treatment at each level of Transfer
treatment_transfer_emmeans <- emmeans(model, ~ Treatment | Transfer)

# Pairwise comparisons between treatments within each level of Transfer
pairwise_comparisons_interaction <- contrast(treatment_transfer_emmeans, method = "pairwise")

# Step 3: Output results
print(pairwise_comparisons_interaction)

# Optional: Adjust for multiple comparisons using Tukey's method
pairwise_comparisons_interaction_tukey <- contrast(treatment_transfer_emmeans, method = "pairwise", adjust = "tukey")
print(pairwise_comparisons_interaction_tukey)





```

```{r}
library(MASS)
library(emmeans)
# Convert Treatment to a factor and reassign it
galdata$Treatment <- as.factor(galdata$Treatment)

# Relevel the factor to make "YPD" the reference level
galdata$Treatment <- relevel(galdata$Treatment, ref = "YPD")


GALsize <- glmmPQL(Number ~ Treatment * Transfer, 
                   random = ~ 1 | Replicate, 
                   family = quasipoisson(link = "log"), 
                   data = galdata)
summary(GALsize)


# Compute estimated marginal means for Treatment within each level of Transfer
emm_interaction <- emmeans(GALsize, ~ Treatment | Transfer)

# Perform pairwise comparisons within each level of Transfer
interaction_comparisons <- contrast(emm_interaction, method = "pairwise", adjust = "bonferroni")

# Display results
summary(interaction_comparisons)

```

```{r}

# Convert Treatment to an unordered factor
Data$Treatment <- as.factor(Data$Treatment)

# Relevel Treatment to set YPD as the baseline
Data$Treatment <- relevel(Data$Treatment, ref = "YPD")

# Fit the model with the new baseline
model <- lm(OD ~ Date * Treatment, data = Data)

# Display the summary of the model
summary(model)

# Verify that YPD is now the baseline treatment
print(levels(Data$Treatment))  # Check the order of levels

# Generate predictions from the model
# Ensure missing data is properly handled in the model
model <- lm(OD ~ Date * Treatment, data = Data, na.action = na.exclude)

# Generate predictions for the model, ensuring alignment with the original data
Data$Predicted <- predict(model)

# Plot actual data points and model predictions
library(ggplot2)

ggplot(Data, aes(x = Date)) +
  geom_point(aes(y = OD, color = Treatment), alpha = 0.6, size = 2) +  # Actual data
  geom_line(aes(y = Predicted, color = Treatment), size = 1) +  # Model predictions
  labs(
    title = "Model Predictions vs. Actual Data",
    x = "Date (Days)",
    y = "OD"
  ) +
  theme_minimal() +
  theme(legend.title = element_text(face = "bold")) +
  scale_color_brewer(palette = "Set2")







```
